A Polls-of-Polls Forecast for the 2024 United States Presidential Election Through Generalized Linear Model

### 2024-11-04 paper/paper.qmd
code draft for discussion

harris_model <- readRDS(file = here("models/vote_for_Harris_model.rds"))
prediction <- read_csv(file = here("data/02-analysis_data/popular_vote_predictions.csv"), show_col_types = FALSE)
```

@@

To assess model reliability, we examined several key diagnostics. Convergence metrics, such as Rhat values, were close to 1 for all parameters, indicating strong convergence. Additionally, the effective sample size (n_eff) was high across parameters, suggesting low autocorrelation and contributing to model stability. For validation, we conducted out-of-sample testing and calculated the Root Mean Square Error (RMSE) to assess predictive accuracy. This test data, which includes polls not used in training, provides an unbiased estimate of model performance. See more details of model diagnostic here @sec-model-details-diagnostics.

@@

## How has the Forecast Changed Over Time? {#sec-first-point}

@@

## Who is Favoured to Win Each State? {#sec-second-point}

## What are the Closest Races Across States {#sec-third-point}

>>>>

harris_model <- readRDS(file = here("models/vote_for_Harris_model.rds"))
prediction <- read_csv(file = here("data/02-analysis_data/vote_for_Harris_predictions.csv"), show_col_types = FALSE)
```

@@

To assess model reliability, we examined several key diagnostics. Convergence metrics, such as Rhat values, were close to 1 for all parameters, indicating strong convergence. Additionally, the effective sample size (n_eff) was high across parameters, suggesting low autocorrelation and contributing to model stability. For validation, we conducted out-of-sample testing and calculated the Root Mean Square Error (RMSE) to assess predictive accuracy. This test data, which includes polls not used in training, provides an unbiased estimate of model performance. See more details of model diagnostic here: @sec-model-details-diagnostics.

@@

## How has the Forecast Changed Over Time? {#sec-first-point}
```{r}
filtered_prediction <- prediction %>%
  filter(national_poll == 1)

# Plot the graph with both predicted and real percentages
ggplot(filtered_prediction, aes(x = end_date)) +
  geom_line(aes(y = predicted_pct, color = "Predicted Percentage")) +
  geom_line(aes(y = pct, color = "Real Percentage")) +
  labs(
       x = "Date",
       y = "Percentage",
       color = "Legend") +
  theme_minimal()
```

@@

## Which States Prefer Harris? {#sec-second-point}
```{r}
state_data <- prediction %>% filter(national_poll == 0)
# Identify missing data
missing_data_summary <- prediction %>%
  group_by(state) %>%
  summarize(missing_dates = sum(is.na(predicted_pct)),
            total_dates = n())

print(missing_data_summary)

# Plot with points for sparse data
ggplot(prediction, aes(x = end_date, y = predicted_pct, color = state)) +
  geom_line(na.rm = TRUE) +
  geom_point(na.rm = TRUE) +  # Add points to show actual data entries
  facet_wrap(~ state, scales = "free_y") +
  labs(
       x = "Date",
       y = "Predicted Percentage (%)") +
  theme_minimal() +
  theme(legend.position = "none",
        strip.text = element_text(size = 10))  # Increase facet label size

```{r}
# Filter to keep only the most recent entry for each state based on 'end_date'
filtered_data <- prediction %>%
  filter(!is.na(state)) %>%
  arrange(state, end_date) %>%
  group_by(state) %>%
  slice_tail(n = 1) %>%
  ungroup() %>%
  select(state, end_date, predicted_pct)

# Plot the bar chart of predicted percentage for each state with pastel colors
ggplot(filtered_data, aes(x = reorder(state, predicted_pct), y = predicted_pct, fill = state)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(predicted_pct, 1)), hjust = -0.1) + # Display values above bars
  labs(
    x = "Predicted %",
    y = "State"
  ) +
  theme_minimal() +
  coord_flip() + # Flip coordinates for better readability
  scale_fill_manual(values = scales::hue_pal()(nrow(filtered_data))) + # Pastel-like palette
  theme(legend.position = "none") # Remove legend for simplicity
```
